Exploring probability distributions for bivariate temporal granularities

### 2019-11-24 main.Rmd
spell check

Calendar-based graphics[@wang2019tsibble] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.  Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009-pk] with days of the week, month of the year and another sub-daily deconstruction of time. But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time. There can be several classes of time deconstructions, viz. based on arrangement (linear vs. cyclic) or hierarchical order of calendar. Linear time granularities respect the linear progression of time such as hours, days, weeks and months. One of the first attempts of characterising these granularities occur in [@Bettini1998-ed]. The definitions and rules defined are inadequate to reflect periodicities in time. Hence, there is a need to define cyclic time granularities in a different approach, which can be useful in visualizing periodic behavior. Cyclic time granularities can be circular or aperiodic depending on if mappings between granularities are regular or irregular. Examples of circular can be hour of the day, day of the week and that of aperiodic granularities can be day of the month or public holidays. Time deconstructions can also be based on the hierarchical structure of time. For example, hours are nested within days, days within weeks, weeks within months, and so on. Hence, it is possible to construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour. lubridate [@G_Grolemund2011-vm] creates easy access and manipulation of common date-time objects. But most accessor functions are limited to single-order-up granularities.

@@

The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world. Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012. Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning. The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks. It is common to see aggregates of usage [@2012-la] across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate. But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency. We propose that the analysis of smart meter data can be benefitted from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies. However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analysed for different periodicities.

@@

The remainder of the paper is organized as follows. Section \ref{sec:timegrandef} details the background of linear granularities in depth and introduces the characterisation of circular and aperiodic time granularities. Section \ref{sec:synergy} discusses the effect in data structure due to how pairs of cyclic time granularities are related. Section \ref{sec:visualization} discusses the role of plot choices, synergy of time granularities and number of observations to construct an informative and trustworthy visualization. Section \ref{sec:application} discusses how systematic exploration can be carried out for a temporal and non-temporal case. Section \ref{sec:discussion} summarizes this paper and discusses possible future direction.

@@

In a cyclic organization of time, the domain is composed of a set of recurring time values. Hence, any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also suceeds Wednesday). Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.

@@

The mappings of two linear granularities forming a cyclic granularity can be regular or irregular. It will be regular if, for example, they are formed by grouping operation. Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity. However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity. The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities. However, the formalisms would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.

@@

<!-- Q2: Amongst good fielding and bowling - which affect the runs of the subsequent overs more? -->

>>>>

Calendar-based graphics[@wang2019tsibble] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.  Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009-pk] with days of the week, month of the year and another sub-daily deconstruction of time. But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time. There can be several classes of time deconstructions, viz. based on arrangement (linear vs. cyclic) or hierarchical order of calendar. Linear time granularities respect the linear progression of time such as hours, days, weeks and months. One of the first attempts of characterizing these granularities occur in [@Bettini1998-ed]. The definitions and rules defined are inadequate to reflect periodicities in time. Hence, there is a need to define cyclic time granularities in a different approach, which can be useful in visualizing periodic behavior. Cyclic time granularities can be circular or aperiodic depending on if mappings between granularities are regular or irregular. Examples of circular can be hour of the day, day of the week and that of aperiodic granularities can be day of the month or public holidays. Time deconstructions can also be based on the hierarchical structure of time. For example, hours are nested within days, days within weeks, weeks within months, and so on. Hence, it is possible to construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour. lubridate [@G_Grolemund2011-vm] creates easy access and manipulation of common date-time objects. But most accessor functions are limited to single-order-up granularities.

@@

The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world. Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012. Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning. The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks. It is common to see aggregates of usage [@2012-la] across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate. But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency. We propose that the analysis of smart meter data can be benefited from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies. However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analysed for different periodicities.

@@

The remainder of the paper is organized as follows. Section \ref{sec:timegrandef} details the background of linear granularities in depth and introduces the characterization of circular and aperiodic time granularities. Section \ref{sec:synergy} discusses the effect in data structure due to how pairs of cyclic time granularities are related. Section \ref{sec:visualization} discusses the role of plot choices, synergy of time granularities and number of observations to construct an informative and trustworthy visualization. Section \ref{sec:application} discusses how systematic exploration can be carried out for a temporal and non-temporal case. Section \ref{sec:discussion} summarizes this paper and discusses possible future direction.

@@

In a cyclic organization of time, the domain is composed of a set of recurring time values. Hence, any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday). Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.

@@

The mappings of two linear granularities forming a cyclic granularity can be regular or irregular. It will be regular if, for example, they are formed by grouping operation. Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity. However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity. The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities. However, the formalism would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.

@@

<!-- Q2: Among good fielding and bowling - which affect the runs of the subsequent overs more? -->
